From DSL to HPC Component-Based Runtime:\\ A Multi-Stencil DSL Case Study

### 2015-08-05 WOLFHPC15/article/component.tex
back to the future

\begin{figure}[t]
\begin{center}
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[provide] (Dp) at (-1,0) {};
   \node (Ds) at (-1.5,0) {start};
   \node[component] (D) at (0,0) {$Driver$};
   \node[use,right=1.5cm of D] (Du1) {};
   \node[use,below=0.8cm of D] (Du2) {};
   \node[use,right=1.7cm of Du2] (Du3) {$m$};

   \node[provide,below=0.15 of Du2] (Tp) {};
   \node[component,below=0.8cm of Tp] (T) {$Time$};
   \node[use,below=0.8cm of T] (Tu) {};
   \node[left=0.8cm of T] (tt) {$T$};

   \node[provide,below=0.15 of Tu] (Sp) {};
   \node[component,below=0.8cm of Sp] (S) {$Scheduling$};
   \node[use,right=1.5cm of S] (Su) {};
   \node[use,right=1.5cm of Tu] (Su2) {};
   \node[left=1.5cm of S] (gamma) {$\Gamma_{tsp}$};

   \node[provide,right=0.15 of Su] (Cp) {};
   \node[component,right=1.5cm of Cp] (C) {$Kernels$};
   \node[use,above=0.8cm of C] (Cu) {};
   \node[right=0.2cm of Cu] (star) {$*$};

   \node[provide,right=0.2 of Du3] (Datap1) {};
   \node[provide,above=0.15 of Cu] (Datap2) {};
   \node[provide,right=0.15 of Su2] (Datap3) {};
   \node[component,above=0.8cm of Datap2] (Data) {$Data$};
   \node[use,above=0.8cm of Data] (Datau) {};
   \node[right=1cm of Data] (delta) {$\Delta,\mathcal{D}$};

@@

  \path[-]
    (Dp)     edge node {} (D)
    (D)      edge node {} (Du1)
             edge node {} (Du2)
             edge node {} (Du3)
    (DDSp1)  edge node {} (DDS)
    (T)      edge node {} (Tu)
    (Tp)     edge node {} (T)
    (Sp)     edge node {} (S)
    (S)      edge node {} (Su)
             edge node {} (Su2)
    (Cp)     edge node {} (C)
    (C)      edge node {} (Cu)
    (Datap1) edge node {} (Data)
    (Datap2) edge node {} (Data)
    (Datap3) edge node {} (Data)
    (Data)   edge node {} (Datau)
    (DDSp2)  edge node {} (DDS);
\end{tikzpicture}
\vspace*{.5em}
\caption{MSCAC Component Assembly Back-End Overview.}
\label{fig:mscac:assembly}
\end{center}
\end{figure}

As described in Section~\ref{sect:mscac}, the compiler generates five independent parts in a component assembly in addition to a static part.
This is represented on Figure~\ref{fig:mscac:assembly} where each rounded box represents one or more component instances.
\texttt{Driver} is static and drives the whole application while \texttt{Kernels} contains code directly written by the user.
The four other parts are generated based on the content of sections of the MSL program: \texttt{Scheduling} from $\Gamma$, \emph{DDS} from $\mathcal{M}$ and $\mathcal{E}$, \emph{Data} from $\Delta$ and $\mathcal{D}$, and \emph{Time} from $T$.

\texttt{DDS} manages the structure of the mesh, a single instance is generated as MSL only supports a single mesh for now.
The back-end presented in this paper uses the distributed data structure proposed in SkelGIS library~\cite{CPE:CPE3494}.
\texttt{DDS} is used by \texttt{Data} containing one component instance for each data element in the MSL program.

\texttt{Time} is parametrized by the the \texttt{iteration} value from MSL and calls the \texttt{Scheduling} part as many times as required.
\texttt{Scheduling} contains the static scheduling computed by MSC, \ie $\Gamma_{tsp}$.
The series-parallel tree is encoded using three specific components which manage the $P$, $S$, and $sync$ operations of $\Gamma_{tsp}$ as explained hereafter.
At the leafs of the tree, the components instances make call either to \texttt{Data} for synchronization of to \texttt{Kernels} for computation.

\texttt{Kernels} contains code written by the user to implement the actual computation.
As presented on Figure~\ref{fig:k}, these components provide a single port used to launch the computation and expose one use port for each data element manipulated in the computation.
We denote the fact that they may have several use port by adding a star on the use port.
Each of these use ports is connected to the matching component instance in \texttt{Data}.
We denote this kind of component K.

As a result of this approach, each part of the generated code is rather independent.
Changing the main loop to use a convergence criteria rather than a fixed number of iterations would only require changing the \texttt{Time} component.
Similarly, changing the approach used for scheduling would only impact the \texttt{Scheduling} components.
The technology used for data parallelism can also be changed by replacing the \texttt{DDS} and \texttt{Data} components; as a matter of fact work is in progress to propose an alternative based on Global Arrays~\cite{Nieplocha:2006:AAP:1125980.1125985} instead of SkelGIS.
In this case however, the \texttt{Kernels} components written by the user have to be adapted to the new interface, but this is the only impact.

>>>>

\begin{figure}[t]
\begin{center}
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[component] (D) at (0,0) {$Driver$};
   \node[provide] (Dp) at (-1,0) {};
   \node (Ds) at (-1.5,0) {start};
   \node[use,right=1.5cm of D] (Du1) {};
   \node[use,below=0.8cm of D] (Du2) {};

   \node[provide,below=0.15 of Du2] (DAp) {};
   \node[component,below=0.8cm of DAp] (DA) {$DriverApp$};
   \node[use,right=1.7cm of DA] (DAu1) {$m$};
   \node[use,below=0.8cm of DA] (DAu2) {};

   \node[provide,below=0.15 of DAu2] (Tp) {};
   \node[component,below=0.8cm of Tp] (T) {$Time$};
   \node[use,right=1cm of T] (Tu) {};
   \node[left=0.8cm of T] (tt) {$T$};

   \node[provide,right=0.15 of Tu] (Cp) {};
   \node[component,right=2cm of Cp] (C) {$Computations$};
   \node[use,above=0.8cm of C] (Cu) {};
   \node[right=0.2cm of Cu] (star) {$*$};
   \node[right=1.5cm of C] (gamma) {$\Gamma$};

   \node[provide,right=0.2 of DAu1] (Datap1) {};
   \node[provide,above=0.15 of Cu] (Datap2) {};
   \node[component,above=0.8cm of Datap2] (Data) {$Data$};
   \node[use,above=0.8cm of Data] (Datau) {};
   \node[right=1cm of Data] (delta) {$\Delta,\mathcal{D}$};

@@

  \path[-]
    (Dp) edge node {} (D)
    (D) edge node {} (Du1)
        edge node {} (Du2)
    (DDSp1) edge node {} (DDS)
    (DA) edge node {} (DAu1)
           edge node {} (DAu2)
    (DAp) edge node {} (DA)
    (Tp) edge node {} (T)
    (T)  edge node {} (Tu)
    (Cp) edge node {} (C)
    (C) edge node {} (Cu)
    (Datap2) edge node {} (Data)
    (Datap1) edge node {} (Data)
    (Data) edge node {} (Datau)
    (DDSp2) edge node {} (DDS);
\end{tikzpicture}
\vspace*{.5em}
\caption{MSCAC Component Assembly Back-End Overview.}
\label{fig:mscac:assembly}
\end{center}
\end{figure}

As described in Section~\ref{sect:mscac}, the compiler generates five independent parts in a component assembly in addition to a static part.
This is represented on Figure~\ref{fig:mscac:assembly} where each rounded box represents one or more component instances.
\texttt{Driver} \& \texttt{DriverApp} are static and drive the whole application.
The four other parts are generated based on the content of sections of the MSL program: \texttt{Computations} from $\Gamma$ (with code provided by the user), \emph{DDS} from $\mathcal{M}$ and $\mathcal{E}$, \emph{Data} from $\Delta$ and $\mathcal{D}$, and \emph{Time} from $T$.

\texttt{DDS} manages the structure of the mesh, a single instance is generated as MSL only supports a single mesh for now.
The back-end presented in this paper uses the distributed data structure proposed in SkelGIS library~\cite{CPE:CPE3494}.
\texttt{DDS} is used by \texttt{Data} containing one component instance for each data element in the MSL program.
\texttt{Time} is parametrized by the the \texttt{iteration} value from MSL and calls the \texttt{Scheduling} part as many times as required.

\texttt{Computations} contains the static scheduling computed by MSC, \ie $\Gamma_{tsp}$.
The series-parallel tree is encoded using three specific components which manage the $P$, $S$, and $sync$ operations of $\Gamma_{tsp}$ as explained hereafter.
At the leafs of the tree, the components instances make call either to \texttt{Data} for synchronization of user-provided components for the computations.
As presented on Figure~\ref{fig:k}, these components provide a single port used to launch the computation and expose one use port for each data element manipulated in the computation.
We denote the fact that they may have several use port by adding a star on the use port.
Each of these use ports is connected to the matching component instance in \texttt{Data}.
We denote this kind of component K.

As a result of this approach, each part of the generated code is rather independent.
Changing the main loop to use a convergence criteria rather than a fixed number of iterations would only require changing the \texttt{Time} component.
Similarly, changing the approach used for scheduling would only impact the code generated in \texttt{Computations}.
The technology used for data parallelism can also be changed by replacing the \texttt{DDS} and \texttt{Data} components; as a matter of fact work is in progress to propose an alternative based on Global Arrays~\cite{Nieplocha:2006:AAP:1125980.1125985} instead of SkelGIS.
In this case however, the \texttt{Computations} part might have to be altered to match the new interface of \texttt{Data}, but this is the only impact.
